NoTCP: steganographic tricks to bypass middleboxes

### 2014-07-16 paper/paper.tex
another iteration of comments from toby, version given to Anil

\begin{figure}[t!]
\centering
\small{
\includegraphics[width=.99\columnwidth]{figs/tcp-header}
\vspace{-5mm}
\caption{TCP Header with overloaded fields and extra control information (Checksum Correction / Control Stream) in payload}
\label{fig:header}
}
\vspace{-4mm}
\end{figure}

Figure~\ref{fig:header} shows the TCP header with fields that we propose for reuse and define later in the paper. However, the structure of TCP header is very simple and gives few choices for hiding extra control information:
\begin{itemize}
    \item Setting header fields without the corresponding flags so they have no meaning to legacy implementations.
    \item Assigning one or more of the reserved field bits.
    \item Overloading the semantics for valid header fields.
\end{itemize}

\subsubsection*{Unset Flags}
\label{sec:unset}

Only 2 header fields have their semantics set by corresponding flags. These are the \emph{Acknowledgment Number} and \emph{Urgent Pointer} fields. The former must also always be set after the initial SYN packet.
%The only header fields that may not have a meaning depending on the corresponding flag are \emph{Urgent Pointer} and \emph{Acknowledgment Number}. The latter must always be set after the initial SYN packet.

The general recommendation is to not use the urgent mechanism~\cite{Gont:2011vi}, therefore it could be reused for other purposes if both endpoints understand the new semantics. On the other hand, it is common practice by Network Intrusion Detection Systems (NIDS) on enterprise networks to reset the field as it makes it difficult to track application-layer data~\cite{seolma}. Importantly, the field becomes \emph{Non-Urgent}~\cite{Kuhlewind:2014vd} when URG is not set -- it has no meaning for legacy implementations and so can be redefined.

@@

3 TCP Reserved bits remain unallocated. One further bit is only allocated to the experimental ECN nonce\cite{Ely:uc}. Despite their limited size, there are proposals that make use of new reserved bits due to implementation simplicity and non-interference with legacy implementations. One suggestion for reducing Web latency~\cite{Flach:2013uy} by mitigating the effects of a small number of lost packets for short connections is to send multiple copies of packets. The proposal deliberately sends copies, but uses a reserved flag to explicitly mark solution avoids triggering fast retransmit by such duplicate acknowledgments by marking them as duplicates in a reserved flag.

@@

The remaining choice is to encode information within otherwise valid header fields. It involves generating packets with specific values for initial sequence numbers, window size or checksums. Arguably it is an acceptable choice if deployable in today's networks. Indeed SYN Cookies already do this~\cite{Eddy:2007to}

Initial sequence numbers must be hard to predict as they provide a measure of security against TCP injection attacks. Adding information to the field to be easily decoded by the receiver makes it predictable and weakens already problematic security~\cite{Bellovin:uz,Qian:2012wb}. Some stateful middleboxes are known to interfere with sequence numbers to improve security instead making it worse~\cite{Qian:2012bj}. Stronger security measures such as TCPCrypt~\cite{Mazieres:uz} are better suited for connection protection.

Window size is potentially a better candidate, especially if used during connection setup phase since data is not typically transmitted during the handshake. Using lower-order bits for control information would prevent flow control from operating properly, but higher-order bits could instead be given up when window scaling option is used to compensate for the decrease of advertised window size. Using the field risks interfering both with TCP flow control mechanism as well as window size optimizations by the network~\cite{Kopparty:2002ht,Chakravorty:2003dm} performed \eg by Citrix ByteMobile Proxy~\cite{Ha:2006td}.

There are multiple benefits to using the checksum field. Firstly, it is only used for error detection. It may be unnecessary when reliable error detection is done at the layer below, however link-level checks have been shown to be unreliable~\cite{Stone:2000fc}. Instead, it may be replaced with potentially stronger mechanisms \eg MPTCP DSS checksum and TCPCrypt Message Authentication Code) or left for the upper layer.

Secondly, checksum operations are efficient as they only involve one's complement addition. We show later in the paper (Section~\ref{sec:protocol}) how checksums can be generated in a way that the intended value is recovered by the receiver after traversing multiple layers of Network Address Translation (NAT).

@@

To measure current network behavior we have developed an Android application distributed through the \emph{Google Play Store}\footnote{Link omitted to anonymize submission} to collect data from volunteers. It contains a simplified implementation of TCP which generates packet sequences that appear valid to network middleboxes: three-way handshake, data exchange and connection teardown. It allows us to modify various header fields to our purposes and check the values received by the other endpoint. We control both endpoints of each connection, therefore we can detect changes from sent values received by the other endpoint.

@@

%Transparency of proxies may also depend on higher-level protocols, therefore we checked for HTTP proxies using techniques proposed in the work, including checking content and header changes, different IP addresses seen for HTTP or non-HTTP traffic and whether non-HTTP traffic on HTTP port is filtered.

\subsection{Limitations}

Potential limitations of our methodology include that we do not consider packet splitting and coalescing. By exchanging information within packet headers and using small packets we avoid packet splitting. We also prevent splitting during data exchanging by not send more than a few bytes of data. Coalescing is prevented by only exchanging a single data message each way. For the same reason we do not take congestion control into account.

@@

\begin{table*}[t]
{\small
\begin{center}
\begin{tabular}{| l | >{\centering\arraybackslash}m{1.3cm} | >{\centering\arraybackslash}m{1.6cm} | >{\centering\arraybackslash}m{2.4cm} | >{\centering\arraybackslash}m{2.4cm} | >{\centering\arraybackslash}m{2.3cm} | >{\centering\arraybackslash}m{1.5cm} | >{\centering\arraybackslash}m{1.5cm} | >{\centering\arraybackslash}m{1.cm} | }
\hline
    \textbf{Net ID} & \textbf{Port-specific} & \textbf{Validate checksum} & \textbf{Drop SYN-ACK data} & \textbf{Normalize SYN Ack number}  & \textbf{Normalize Urgent Pointer} & \textbf{Normalize Reserved} & \textbf{Remap Sequence} \\ \hline \hline
    % Univ Helsinki WiFi
    WiFi edu 1      &                        & \checkmark                 &                        &                         &                        &                             &                     \\ \hline
    % Eduroam
    WiFi edu 2      &                        &                            & \checkmark             &                         & \checkmark             &                             & \checkmark          \\ \hline
    % wgb
    WiFi pub 1      &                        &                            &                        &                         &                        &                             &                     \\ \hline
    % MKSW germany (carlos)
    WiFi pub 2      &                        & \checkmark                 & \checkmark             &                         &                        &                             &                     \\ \hline
    % Virgin home
    WiFi res 1      &                        & \checkmark                 &                        &                         &                        &                             &                     \\ \hline
    % TEO Lt
    WiFi res 2      &                        & \checkmark                 & \checkmark             &                         &                        &                             &                     \\ \hline
    % Telefonica Free
    WiFi ent 1      &                        & \checkmark                 & \checkmark             &                         &                        &                             &                     \\ \hline
    \hline
    % Finland internet.saunalahti
    Cellular 1      &                        &                            &                        &                         &                        &                             &                     \\ \hline
    % GiffGaff
    Cellular 2      &  443, 993              & *                          & *                      & *                       & *                      & *                           & *                   \\ \hline
    % E-Plus germany
    Cellular 3      &                        &                            &                        &                         & \checkmark             &                             & \checkmark          \\ \hline
    % EE/T-Mobile
    Cellular 4      &               80, 443  & *                          & *                      & *                       & *                      & *                           & *                   \\ \hline
    % Three UK
    Cellular 5      &                        & \checkmark                 &                        &                         &                        &                             &                     \\ \hline
    % Omnitel LT
    Cellular 6      &                        & \checkmark                 &                        &                         &                        &                             &                     \\ \hline
\end{tabular}
\end{center}
}
\vspace{-4mm}
\caption{Network behavior observed through tests generating custom TCP packets. * means different cases observed based on port numbers.}
\label{tab:networks}
\vspace{-4mm}
\end{table*}

For our preliminary study we have focused on mobile networks including public, residential, university and enterprise WiFi networks as well as 6 cellular networks in 4 different countries. We present our results in Table~\ref{tab:networks}.

The results include networks that did not filter any modifications (only one WiFi and two cellular nets), and others that interfered with our traffic. The interference we tested for includes resetting fields without corresponding flags set (\emph{Acknowledgment Number} and \emph{Urgent Pointer}) or dropping them, discarding packets with invalid checksums, and resetting reserved flags. Only one network provided clients with a global IP (WiFi edu 2) and we also tested whether NATs rewrite sequence numbers in addition to addresses and port numbers.

As discussed in Section~\ref{sec:unset}, we can squeeze in a few extra bits of information by using certain header fields when the corresponding flags are not set. None of our tested networks filtered the Acknowledgment Number field for the initial SYN packet on non-proxied paths and only two networks discarded packets with the Urgent Pointer set without the URG flag.

We also tested middlebox interaction with reserved header bits (Section~\ref{sec:reserved}). We separated our tests to handshake and data exchange parts as well as set each bit individually, however we did not observe any differences. All networks appear to pass reserved bits on non-proxied connections.

@@

We confirmed the middlebox on Cellular 2 to be a ByteMobile proxy. It replaces the original connections with new ones, adding options not present in original packets, changing window size and resetting other fields to different values to optimize TCP for cellular networks~\cite{Ha:2006td}. Such behavior makes deployment of any protocol modifications or extensions difficult -- we verified that TCPCrypt fails on this network. We do not currently have an estimate of how widespread ByteMobile proxies are as we only observed them on one network.

In the case of \emph{Cellular 4}, traffic was only modified when going to ports 80 and 443. The network likely optimizes HTTP traffic and drops any non-HTTP traffic. Generally port-specific behavior appears related to proxies.

@@

Proxies potentially change every aspect of the original connection, rewrite or filter header values they do not recognize and hiding any information within the packet becomes difficult. We checked for them in our tested networks on a number of popular port numbers:
21, 22, 25, 80, 110, 135, 139, 143, 161, 443, 445, 465, 585, 587, 993, 995, 1194, 1723, 5060, 6881, 9001.

\begin{table}[t!]
{\small
\begin{center}
\begin{tabular}{| l | >{\centering\arraybackslash}m{1.8cm} | >{\centering\arraybackslash}m{3.7cm} |}
\hline
    \textbf{Net ID} & \textbf{Radio Technology} & \textbf{Proxied Ports} \\ \hline \hline
    Cellular 2      & HSPA                      & 21, 22, 25, 80, 110, 135, 143, 161, 465, 585, 587, 995, 1194, 1723, 5060, 6881, 9001 \\ \hline
    Cellular 4      & GPRS                      & 80, 110, 143 \\ \hline
    Cellular 4      & HSPA                      & 80, 110, 143 \\ \hline
    Cellular 4      & LTE                       & 80, 110, 143, 443, 993, 995 \\ \hline
    % T-Mobile, part of EE
    Cellular 4*     & HSPA                     & 25, 80, 110, 143 \\ \hline
    % Virgin on T-Mobile
    Cellular 4**    & HSPA                    & 25, 80, 110, 143 \\ \hline
\end{tabular}
\end{center}
}
\vspace{-4mm}
\caption{Non-transparent proxies and their proxied port numbers by radio technology. */** - see text for details.}
\label{tab:proxies}
\vspace{-4mm}
\end{table}

We summarize results in Table~\ref{tab:proxies}. 3G and 4G networks use different IP core networks, therefore we split the results by radio technology. We did not detect proxies on Cellular 1 and Cellular 3 networks and did not have data about Cellular 6, but results of our TCP testsuite did not indicate any proxies. Cellular 2 proxies almost all of our tested ports except for 443 and 993.

Cellular 4 shows non-uniform behavior across the network. It is not obvious why the network proxies connections to ports commonly used for encrypted communications. Currently we have only recorded 1 such case for HTTPS. The differences between the other two cases (* and **) seem related to forming the operator from two merged ones. Their infrastructure may have not been completely consolidated: \emph{whois} records and device-reported network names are different. Case \textbf{*} shows the old operator name and \textbf{**} reports the name of a virtual network operator (MVNO). MVNOs buy network services from operators at wholesale rates and typically have little infrastructure of their own, and expect to see the same behavior as that of the underlying network.

Estimating proxying from \emph{whois} records alone is not safe: they do not necessarily match the operator names and we saw the same address ranges used with all device-reported network names for HSPA and LTE radios. Reliable detection needs active testing, potentially combined with caching for repeated connections.

%Proxying is protocol-dependent in some cases: looking at port 80, Cellular 4\*\* modified HTTP headers, performed transcoding and reordered header fields, but only modified non-HTTP traffic for one of the two sessions. While some networks did not interfere with our tests, they do not appear to be completely free of proxies. On Cellular 1 (15 sessions) two sessions show DNS proxies (over UDP) and on Cellular 5 two sessions (out of 47) had non-HTTP traffic filtered on port 80. Finally, Cellular 2 and 4 both reorder HTTP headers and perform transcoding, but Cellular 4 does not always allow non-HTTP traffic.

%    an_op_name    | an_cell_net_type | an_country | port |    ip_global (simplified to network address. Included the owner according to whois
% db)
% -----------------+------------------+------------+------+-----------------
%  EE              | GPRS             | gb         |   80 | 31.72/19 (EE)
%  EE              | GPRS             | gb         |  110 | 31.72/19 (EE)
%  EE              | GPRS             | gb         |  143 | 31.72/19 (EE) 
%  EE              | HSPA             | gb         |   25 | 31.96/21 (EE), 149.254/16 (T-Mobile UK), 178.96/12 (EE)
%  EE              | HSPA             | gb         |   80 | 31.80/21 (EE), 31.96/21 (EE), 149.254/16 (T-Mobile UK), 178.96/12 (EE), 213.205.224/12 (EE)
%  EE              | HSPA             | gb         |  110 | 31.80/21 (EE), 31.96/21 (EE), 149.254/16 (T-Mobile UK), 178.96/12 (EE), 213.205.224/12 (EE)
%  EE              | HSPA             | gb         |  143 | 31.80/21 (EE), 31.96/21 (EE), 149.254/16 (T-Mobile UK), 178.96/12 (EE), 213.205.224/12 (EE)
%  EE              | LTE              | gb         |   80 | 31.96/21 (EE), 89.192/16 (EE), 213.205.224/12 (EE)
%  EE              | LTE              | gb         |  110 | 31.96/21 (EE), 89.192/16 (EE), 213.205.224/12 (EE)
%  EE              | LTE              | gb         |  143 | 31.96/21 (EE), 89.192/16 (EE), 213.205.224/12 (EE)
%  EE              | LTE              | gb         |  443 | 213.205.224/12 (EE)
%  EE              | LTE              | gb         |  993 | 31.96/21 (EE), 89.192/16 (EE), 213.205.224/12 (EE)
%  EE              | LTE              | gb         |  995 | 31.96/21 (EE), 89.192/16 (EE), 213.205.224/12 (EE)
%  T-Mobile        | HSPA             | gb         |   25 | 31.96/21 (EE), 149.254/16 (T-Mobile)
%  T-Mobile        | HSPA             | gb         |   80 | 31.96/21 (EE), 149.254/16 (T-Mobile)
%  T-Mobile        | HSPA             | gb         |  110 | 31.96/21 (EE), 149.254/16 (T-Mobile)
%  T-Mobile        | HSPA             | gb         |  143 | 31.96/21 (EE), 149.254/16 (T-Mobile)
%  Virgin          | HSPA             | gb         |   25 | 149.254/16 (T-Mobile)
%  Virgin          | HSPA             | gb         |   80 | 149.254/16 (T-Mobile)
%  Virgin          | HSPA             | gb         |  110 | 149.254/16 (T-Mobile)
%  Virgin          | HSPA             | gb         |  143 | 149.254/16 (T-Mobile)
%  giffgaff        | HSPA             | gb         |   21 | 82.132/16 (O2)
%  giffgaff        | HSPA             | gb         |   22 | 82.132/16 (O2)
%  giffgaff        | HSPA             | gb         |   25 | 82.132/16 (O2)
%  giffgaff        | HSPA             | gb         |   80 | 82.132/16 (O2)
%  giffgaff        | HSPA             | gb         |  110 | 82.132/16 (O2)
%  giffgaff        | HSPA             | gb         |  135 | 82.132/16 (O2)
%  giffgaff        | HSPA             | gb         |  143 | 82.132/16 (O2)
%  giffgaff        | HSPA             | gb         |  161 | 82.132/16 (O2)
%  giffgaff        | HSPA             | gb         |  465 | 82.132/16 (O2)
%  giffgaff        | HSPA             | gb         |  585 | 82.132/16 (O2)
%  giffgaff        | HSPA             | gb         |  587 | 82.132/16 (O2)
%  giffgaff        | HSPA             | gb         |  995 | 82.132/16 (O2)
%  giffgaff        | HSPA             | gb         | 1194 | 82.132/16 (O2)
%  giffgaff        | HSPA             | gb         | 1723 | 82.132/16 (O2)
%  giffgaff        | HSPA             | gb         | 5060 | 82.132/16 (O2)
%  giffgaff        | HSPA             | gb         | 6881 | 82.132/16 (O2)
%  giffgaff        | HSPA             | gb         | 9001 | 82.132/16 (O2)

@@

What is common to all candidate protocol changes described above is that they are very limited in size and therefore only sufficient to exchange an opcode. Other protocol details must instead be exchanged either out of band, implicit to the opcode, or explicitly by redefining the meaning of part of the data payload or previously exchanged information. Here we propose one way of exchanging control information as an opcode to bootstrap a full control stream.

@@

A surprising result in our tests is that the \emph{Acknowledgment Number} field without the ACK flag set is passed on all non-proxied connections. As it is a 32-bit value we can allocate part of it (\eg 8 or 16 bits) to signal that the rest of the field is significant. The exact allocation is subject to further discussion, but we suggest using the first 16 bits as \emph{Extension Space} and the others as $Opcode$ (Figure~\ref{fig:header}), carrying the desired opcode. For simultaneous open we can use the same mechanism in both directions.

For acknowledging extensions in SYN-ACK packet we use the Urgent Pointer and Checksum fields. Our tests indicate that neither choice is guaranteed to succeed, therefore we combine the two by redefining their values when SYN=1. We define the Urgent Pointer to be $Opcode_U$ when URG=0 and Checksum to carry a meaning when Urgent Pointer = 0 to get around the cases where one is blocked but the other one is not. Across all networks we tested in our preliminary study, at least one of the choices succeeds.

Checksum value can not be used directly as an opcode since it changes at every level of NAT. The traditional NAT does a simple recalculation of the checksum~\cite{Egevang:tu}: subtracts the original source address and port number (in some cases also the sequence number) and adds the new values. The received Checksum after \(n\) NAT hops is therefore:

\vspace{-2mm}
\begin{align*}
Checksum_2 & = Checksum_1 - header_1 + header_2 \\
Checksum_i & = Checksum_{i-1} - header_{i-1} + header_i \\
 ... \\
Checksum_n & = Checksum_1 - header_1 + header_n
\end{align*}

@@

\vspace{-2mm}
\begin{align*}
Checksum_{target} & = Opcode + header_1 \\
Checksum_n = Opcode_C & = Opcode + header_n
\end{align*}

@@

For our prototype we assign the first 16 bits of the payload as \emph{Checksum Correction} to force a specific checksum to maintain validity of a packet. The choice is in line with other extensions which reuse part of data payload~\cite{Mazieres:uz,Bonaventure:wx}. This Checksum Correction field is not delivered to the application. TLV-Encoded (Type-Length-Value) Control Stream can be present in data part~\cite{Bonaventure:wx} of the SYN-ACK packet after the Checksum Correction field.

Sending data with the SYN-ACK packet works on all tested cellular networks, but not on most WiFi. For the above header modifications the information either reaches the destination as intended, or the packet is dropped. Finally, on all of our tested networks at least one case succeeds as long as there is no non-transparent proxy for the chosen destination port.

@@

We use the embedded opcodes in the handshake to exchange sufficient control information to bootstrap a complete control stream. We extend the three-way handshake with fields shown in Figure~\ref{fig:header}:

@@

\begin{figure}[t!]
\centering
\includegraphics[width=.9\columnwidth]{figs/handshake}
% \vspace{-2mm}
\caption{Example handshake with extension negotiation}
\label{fig:handshake}
\vspace{-4mm}
\end{figure}

@@

We witnessed a case of packet splitting where an ACK segment is split to one only acknowledging the received data and another other one with the actual payload. We did not observe a case where a middlebox both passes payload in SYN-ACK packet and splits ACK segments. If the handshake succeeds we need to take a little extra care with the fallback mechanism -- we need to wait for an ACK segment with payload before falling back to vanilla TCP. If the passive opener receives data it can not decode as valid TLV-encoded control information, it falls back to vanilla TCP.

@@

%Covert channels are typically low bitrate in order to make it difficult to detect them. Application-specific ones, especially based on media transcoding can achieve rates from 270 b/s using $\mu TP$ torrent protocol, to 2 kbit/s by encoding data in silent Skype voice call periods, to megabits in wireless environments by inserting hidden data into the padding of frames~\cite{Zielinska:2014fn}.

Our control channel does not need to protection from steganalysis as it is only used to avoid middlebox interference with end-to-end control information exchange. Bitrate of the channel only using packet headers (\ie without redefining part of packet payload for control information) is therefore bounded by the size of the TCP header (20 bytes without options), the round trip time (RTT) and the number of packets in flight per RTT. Depending on fields usable for a particular connection as inferred during the handshake, we can use between 16 (either $Opcode_C$ or $Opcode_U$) and 35 bits when using both fields and reserved flags.

>>>>

\begin{figure}[t!]
\centering
\small{
\includegraphics[width=\columnwidth]{figs/tcp-header}
% \vspace{-5mm}
\caption{TCP Header with overloaded fields and extra control information (Checksum Correction / Control Stream) in payload}
\label{fig:header}
}
% \vspace{-4mm}
\end{figure}

Figure~\ref{fig:header} shows the TCP header with fields that we propose for reuse (defined later in the paper). The structure of the TCP header is very simple and gives few choices for hiding extra control information:
\begin{itemize}
    \item Setting header fields without the corresponding flags so they have no meaning to legacy implementations.
    \item Assigning one or more of the reserved field bits.
    \item Overloading the semantics for valid header fields.
\end{itemize}

\subsubsection*{Flag-dependent semantics}
\label{sec:unset}

Only two header fields have their semantics set by corresponding flags. These are the \emph{Acknowledgment Number} and \emph{Urgent Pointer} fields. The former must also always be set after the initial SYN packet.

The general recommendation is to not use the urgent mechanism~\cite{Gont:2011vi}, therefore it could be reused for other purposes if both endpoints understand the new semantics. On the other hand, Network Intrusion Detection Systems (NIDS) often reset the field as it makes it difficult to track application-layer data~\cite{seolma}. Importantly, the field becomes \emph{Non-Urgent}~\cite{Kuhlewind:2014vd} when URG is not set -- it has no meaning for legacy implementations and so can be redefined.

@@

Three TCP Reserved bits remain unallocated. One further bit is only allocated to the experimental ECN nonce\cite{Ely:uc}. Despite their limited size, there are proposals that make use of new reserved bits due to implementation simplicity and non-interference with legacy implementations. One suggestion for reducing Web latency~\cite{Flach:2013uy} by mitigating the effects of a small number of lost packets for short connections is to send multiple copies of packets. The proposal deliberately sends copies, but uses a reserved flag to explicitly mark solution avoids triggering fast retransmit by such duplicate acknowledgments by marking them as duplicates in a reserved flag.

@@

The remaining choice is to encode information within otherwise valid header fields. It involves generating packets with specific values for \emph{Initial Sequence Numbers}, \emph{Window Size} or \emph{Checksums}. Arguably it is an acceptable choice if deployable in today's networks. Indeed SYN Cookies already do this~\cite{Eddy:2007to}

\emph{Initial Sequence Numbers} must be hard to predict as they provide a measure of security against TCP injection attacks. Adding information to the field to be easily decoded by the receiver makes it more predictable and weakens already problematic security~\cite{Bellovin:uz,Qian:2012wb}. Also some stateful middleboxes are known to interfere with sequence numbers in an attempt to improve security~\cite{Qian:2012bj}. Stronger security measures such as TCPCrypt~\cite{Mazieres:uz} are better suited for connection protection.

\emph{Window Size} is potentially a better candidate, especially if used during connection setup phase since data is not typically transmitted during the handshake. Using lower-order bits for control information would prevent flow control from operating properly, but higher-order bits could instead be given up when the \emph{window scaling} option is used to compensate for the decrease of advertised window size. Using the field risks interfering both with TCP flow control mechanism as well as window size optimizations by the network~\cite{Kopparty:2002ht,Chakravorty:2003dm} performed by \eg Citrix ByteMobile Proxy~\cite{Ha:2006td}.

There are multiple benefits to using the checksum field. Firstly, it is only used for error detection. It may be unnecessary when reliable error detection is done at the layer below, however link-level checks have been shown to be unreliable~\cite{Stone:2000fc}. Instead, it may be replaced with potentially stronger mechanisms \eg MPTCP DSS checksum and TCPCrypt Message Authentication Code or left for the upper layer. Secondly, checksum operations are efficient as they only involve one's complement addition. We show later in the paper (\S\ref{sec:protocol}) how checksums can be generated in a way that the intended value is recovered by the receiver after traversing multiple layers of Network Address Translation (NAT).

@@

To measure current network behavior we have developed an Android application distributed through the \emph{Google Play Store}\footnote{Link omitted to anonymize submission} to collect data from volunteers. It contains a simplified implementation of TCP which generates packet sequences that appear valid to network middleboxes: three-way handshake, data exchange and connection teardown. It allows us to modify various header fields and check the values received by the other endpoint. We control both endpoints of each connection, therefore we can detect changes from sent values received by the other endpoint.

@@

Our methodology avoids packet splitting by only using small packets. Coalescing is prevented by only exchanging a single data message each way. For the same reason we do not take congestion control into account.

@@

\begin{table*}[t]
{\small
\begin{center}
\begin{tabular}{| l | >{\centering\arraybackslash}m{1.3cm} | >{\centering\arraybackslash}m{1.6cm} | >{\centering\arraybackslash}m{2.4cm} | >{\centering\arraybackslash}m{2.4cm} | >{\centering\arraybackslash}m{2.3cm} | >{\centering\arraybackslash}m{1.5cm} | >{\centering\arraybackslash}m{1.5cm} | >{\centering\arraybackslash}m{1.cm} | }
\hline
    \textbf{Net ID} & \textbf{Port-specific} & \textbf{Validate checksum} & \textbf{Drop SYN-ACK data} & \textbf{Normalize SYN Ack number}  & \textbf{Normalize Urgent Pointer} & \textbf{Normalize Reserved} & \textbf{Remap Sequence} \\ \hline \hline
    % Univ Helsinki WiFi
    WiFi edu 1      &                        & \checkmark                 &                        &                         &                        &                             &                     \\ \hline
    % Eduroam
    WiFi edu 2      &                        &                            & \checkmark             &                         & \checkmark             &                             & \checkmark          \\ \hline
    % wgb
    WiFi pub 1      &                        &                            &                        &                         &                        &                             &                     \\ \hline
    % MKSW germany (carlos)
    WiFi pub 2      &                        & \checkmark                 & \checkmark             &                         &                        &                             &                     \\ \hline
    % Virgin home
    WiFi res 1      &                        & \checkmark                 &                        &                         &                        &                             &                     \\ \hline
    % TEO Lt
    WiFi res 2      &                        & \checkmark                 & \checkmark             &                         &                        &                             &                     \\ \hline
    % Telefonica Free
    WiFi ent 1      &                        & \checkmark                 & \checkmark             &                         &                        &                             &                     \\ \hline
    \hline
    % Finland internet.saunalahti
    Cellular 1      &                        &                            &                        &                         &                        &                             &                     \\ \hline
    % GiffGaff
    Cellular 2      &  443, 993              & *                          & *                      & *                       & *                      & *                           & *                   \\ \hline
    % E-Plus germany
    Cellular 3      &                        &                            &                        &                         & \checkmark             &                             & \checkmark          \\ \hline
    % EE/T-Mobile
    Cellular 4      &               80, 443  & *                          & *                      & *                       & *                      & *                           & *                   \\ \hline
    % Three UK
    Cellular 5      &                        & \checkmark                 &                        &                         &                        &                             &                     \\ \hline
    % Omnitel LT
    Cellular 6      &                        & \checkmark                 &                        &                         &                        &                             &                     \\ \hline
\end{tabular}
\end{center}
}
% \vspace{-4mm}
\caption{Network behavior observed through tests generating custom TCP packets. * means different cases observed based on port numbers.}
\label{tab:networks}
% \vspace{-6mm}
\end{table*}

Our preliminary study focused on mobile networks including public, residential, university and enterprise WiFi networks as well as six cellular networks in four different countries. We present our results in Table~\ref{tab:networks}.

The results include networks that did not filter any modifications (only one WiFi and two cellular nets), and others that interfered with our traffic. We looked for dropped packets and reset or altered fields: (\emph{Acknowledgment Number} and \emph{Urgent Pointer}, potentially invalid checksums and reserved flags. Only one network provided clients with a global IP (WiFi edu 2) so we also tested whether NATs rewrite sequence numbers in addition to addresses and port numbers.

As discussed in \S\ref{sec:unset}, we can squeeze in a few extra bits of information by using certain header fields when the corresponding flags are not set. None of our tested networks filtered the Acknowledgment Number field for the initial SYN packet on non-proxied paths and only two networks discarded packets with the Urgent Pointer set without the corresponding URG flag.

We also tested middlebox interaction with reserved header bits (\S\ref{sec:reserved}). We separated our tests to handshake and data exchange parts as well as set each bit individually, however we did not observe any differences. All networks appear to pass reserved bits on non-proxied connections.

@@

We confirmed the presence of a ByteMobile proxy on Cellular 2. It replaces the original connections with new ones, adding options not present in original packets, changing window size and resetting other fields to different values to optimize TCP for cellular networks~\cite{Ha:2006td}. Such behavior makes deployment of any protocol modifications or extensions difficult -- we verified that TCPCrypt fails on this network. We do not currently have an estimate of how widespread ByteMobile proxies are as we only observed them on one network.

In the case of \emph{Cellular 4}, traffic was only modified when going to ports 80 and 443. The network likely optimizes HTTP traffic and drops non-HTTP traffic. Generally port-specific behavior appears related to proxies.

@@

Proxies potentially change every aspect of the original connection, rewrite or filter header values they do not recognize so hiding any information within the packet becomes difficult. We tested for them on a number of popular port numbers:
21, 22, 25, 80, 110, 135, 139, 143, 161, 443, 445, 465, 585, 587, 993, 995, 1194, 1723, 5060, 6881, 9001.

\begin{table}[t!]
{\small
\begin{center}
\begin{tabular}{| l | >{\centering\arraybackslash}m{1.8cm} | >{\centering\arraybackslash}m{3.7cm} |}
\hline
    \textbf{Net ID} & \textbf{Radio Technology} & \textbf{Proxied Ports} \\ \hline \hline
    Cellular 2      & HSPA                      & 21, 22, 25, 80, 110, 135, 143, 161, 465, 585, 587, 995, 1194, 1723, 5060, 6881, 9001 \\ \hline
    Cellular 4      & GPRS                      & 80, 110, 143 \\ \hline
    Cellular 4      & HSPA                      & 80, 110, 143 \\ \hline
    Cellular 4      & LTE                       & 80, 110, 143, 443, 993, 995 \\ \hline
    % T-Mobile, part of EE
    Cellular 4*     & HSPA                     & 25, 80, 110, 143 \\ \hline
    % Virgin on T-Mobile
    Cellular 4**    & HSPA                    & 25, 80, 110, 143 \\ \hline
\end{tabular}
\end{center}
}
% \vspace{-4mm}
\caption{Non-transparent proxies and their proxied port numbers by radio technology. */** - see text for details.}
\label{tab:proxies}
% \vspace{-4mm}
\end{table}

@@

What is common to all candidate protocol changes described above is that they are limited in size and therefore only sufficient to exchange an opcode. Other protocol details must be exchanged either out of band, implicitly to the opcode, or explicitly by redefining the meaning of part of the data payload or previously exchanged information. Here we propose one way of exchanging control information as an opcode to bootstrap a full control stream.

@@

A surprising result in our tests is that the \emph{Acknowledgment Number} field without the ACK flag set is passed on all non-proxied connections. As it is a 32-bit value we can use part of it to signal that the rest of the field is significant. The exact allocation is subject to further discussion, but we suggest using the first 16 bits as \emph{Extension Space} and the others as $Opcode$ (Figure~\ref{fig:header}). For \emph{simultaneous open} we can use the same mechanism in both directions.

For acknowledging extensions in the SYN-ACK packet we use the Urgent Pointer and Checksum fields. Our tests indicate that neither choice is guaranteed to succeed, therefore we combine the two by redefining their values when SYN=1. We define the Urgent Pointer to be $Opcode_U$ when URG=0. If Urgent Pointer is 0, Checksum is the opcode to get around the cases where one is blocked but the other one is not. Across all networks tested in our preliminary study, at least one of the choices succeeds. In the cases when both are usable data can be encoded across them as a single field once the handshake is completed.

Checksum value can not be used directly as an opcode since it changes at every level of NAT. The traditional NAT does a simple recalculation of the checksum~\cite{Egevang:tu}: subtracts the original source address and port number (in some cases also the sequence number) and adds the new values. The Checksum after \(n\) NAT hops is:

% \vspace{-4mm}
\begin{align*}
Checksum_2 & = Checksum_1 - header_1 + header_2 \\
Checksum_i & = Checksum_{i-1} - header_{i-1} + header_i \\
 ... \\
Checksum_n & = Checksum_1 - header_1 + header_n
\end{align*}
% \vspace{-2mm}

@@

% \vspace{-4mm}
\begin{align*}
Checksum_{target} & = Opcode + header_1 \\
Checksum_n = Opcode_C & = Opcode + header_n
\end{align*}
% \vspace{-2mm}

@@

For our prototype we assign the first 16 bits of the payload as \emph{Checksum Correction} to force a specific checksum to maintain validity of a packet. The field value is not considered as part of application data. The choice is in line with other extensions which reuse part of data payload~\cite{Mazieres:uz,Bonaventure:wx}. Sending data with the SYN-ACK packet works on all tested cellular networks, but not on most WiFi.

For the above header modifications the information either reaches the destination as intended, or the packet is dropped. Finally, on all of our tested networks at least one case succeeds as long as there is no non-transparent proxy for the chosen destination port.

@@

We use the embedded opcodes in the handshake to exchange sufficient control information to bootstrap a complete control stream. After the handshake a higher-bandwidth \emph{Control Channel} can be defined by using TLV (Type-Length-Value) encoding to delimit control information within packet payload~\cite{Bonaventure:wx}. The channel is bootstrapped using an extended three-way handshake with fields shown in Figure~\ref{fig:header}:

@@

\begin{figure}[t!]
\centering
\includegraphics[width=.9\columnwidth]{figs/handshake}
% \vspace{-2mm}
\caption{Example handshake with extension negotiation}
\label{fig:handshake}
% \vspace{-4mm}
\end{figure}

@@

We witnessed one case of packet splitting where an ACK segment is split to one only acknowledging the received data and another other one with the actual payload. We did not observe a case where a middlebox both passes payload in SYN-ACK packet and splits ACK segments. If the handshake succeeds we need to take a little extra care with the fallback mechanism -- we need to wait for an ACK segment with payload before falling back to vanilla TCP. If the passive opener receives data it can not decode as valid TLV-encoded control information, it falls back to vanilla TCP.

@@

Our control channel does not need to protection from steganalysis as it is only used to avoid middlebox interference with end-to-end control information exchange. The bitrate of the channel without redefining part of packet payload for control information is bounded by the size of the TCP header (20 bytes without options), the round trip time (RTT) and the number of packets in flight per RTT. Depending on fields usable for a particular connection as inferred during the handshake, we can use between 16 (either $Opcode_C$ or $Opcode_U$) and 35 bits when using both fields and reserved flags.

For an operating window $cwnd$ (in bytes) a TCP flow has $cwnd/MSS$ data packets per RTT, hence that many headers can carry control data. If \emph{window scaling} option is not used, the maximum \emph{receive window} is 65536 bytes. Using the typical HSPA network RTT of about 100ms the maximum data rate is therefore 5 Mbit/s, reasonable for current networks~\cite{Chen:2012vj}. In this case covert channel bitrate is 7.0 to 15.4 kbit/s, depending on the fields used.
